Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-24 overview.tex
Changed all "X order" to "x-order" and "finite-difference" and "finite-volume" to "finite difference" and "finite volume", to ensure consistency.

Four different (magneto)-hydrodynamic methods are implemented in
\enzo: (i) the hydrodynamic-only piecewise parabolic method (PPM)
developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
\citep{1977JCoPh..23..276V} with or without magnetic fields and Dedner-based
divergence cleaning, described in more detail in
\citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
transport staggered MHD scheme \citep{Collins10}, and (iv) the
second-order finite-difference hydrodynamics method described in
\zeus~\citep{Stone92a,Stone92b}.

@@

We begin with the direct-Eulerian PPM implementation.  This is an
explicit, higher-order-accurate version of Godunov's method for ideal
gas dynamics with the spatially third order accurate piecewise parabolic monotonic
interpolation and a nonlinear Riemann solver for shock capturing.  It
advances the hydrodynamic equations in the following steps:
%\begin{enumerate}
% \item 
(i) Construct monotonic parabolic interpolation of cell average data,
for each fluid quantity;
% \item 
(ii) Compute interface states by averaging the parabola over the
domain of dependence for each interface;
% \item 
(iii) Use interface data to solve the Riemann problem;
 %\item 
(iv) Difference the interface fluxes to update the cell average
quantities.
%\end{enumerate}

The PPM implementation does an excellent job of capturing strong
shocks across a few cells.  Multidimensional schemes are built up by
directional splitting and produce a method that is formally second
order-accurate in space and time and which explicitly conserves mass,
linear momentum, and energy \citep{Hawley84, Norman86}.  A variety of
Riemann solvers have been implemented.

@@

\paragraph{Godunov MHD with Constrained Transport (MHD)}
This MHD method is second order in time and space, and preserves the
divergence constraint, $\div \vecB = 0$, to machine precision
through the Constrained Transport (CT) method \citep{Collins10}.  CT, originally
described by \citet{Evans88}, updates the magnetic field with the curl
of an electric field, suitably formulated to preserve $\div
\vecB$.  We employ the CT methods described by \citet{Balsara99} and
\citet{Gardiner05} with the second order hyperbolic
solver of \citet{Li08a} and the divergence free AMR scheme of
\citet{Balsara01}.

\paragraph{Second order finite difference method (HD only)}

Lastly, we briefly describe the \zeus\ method, a finite-difference
algorithm originally used in the \zeus\ code \citep{Stone92a}. Note
that \enzo\ is entirely independent of the \zeus\ code, and only the
hydrodynamical algorithm of \zeus\ has been implemented in \enzo; the
MHD and radiation-hydrodynamics schemes have not. The \zeus\ method
uses a staggered mesh such that velocities are face-centered, while
density and internal energy are cell-centered.  It splits the solution
up into two steps. The first is the so-called source step, in which
the momentum and energy values are updated to reflect the pressure and
gravity forces, including an artificial viscosity required for
stability. The second step, known as the transport step, accounts for
the advection of conserved quantities (mass, momentum, and energy)
across the grid.

>>>>

Four different (magneto)-hydrodynamic methods are implemented in
\enzo: (i) the hydrodynamic-only piecewise parabolic method (PPM)
developed by~\citet{1984JCoPh..54..174C} and extended to cosmology
by~\citet{1995CoPhC..89..149B}; (ii) the MUSCL-like Godunov scheme
\citep{1977JCoPh..23..276V} with or without magnetic fields and Dedner-based
divergence cleaning, described in more detail in
\citet{WangAbelZhang08} and \citet{WangAbel09}; (iii) a constrained
transport staggered MHD scheme \citep{Collins10}, and (iv) the
second-order finite difference hydrodynamics method described in
\zeus~\citep{Stone92a,Stone92b}.

@@

We begin with the direct-Eulerian PPM implementation.  This is an
explicit, higher-order-accurate version of Godunov's method for ideal
gas dynamics with the spatially third-order-accurate piecewise parabolic monotonic
interpolation and a nonlinear Riemann solver for shock capturing.  It
advances the hydrodynamic equations in the following steps:
%\begin{enumerate}
% \item 
(i) Construct monotonic parabolic interpolation of cell average data,
for each fluid quantity;
% \item 
(ii) Compute interface states by averaging the parabola over the
domain of dependence for each interface;
% \item 
(iii) Use interface data to solve the Riemann problem;
 %\item 
(iv) Difference the interface fluxes to update the cell average
quantities.
%\end{enumerate}

The PPM implementation does an excellent job of capturing strong
shocks across a few cells.  Multidimensional schemes are built up by
directional splitting and produce a method that is formally second-order-accurate in space and time and which explicitly conserves mass,
linear momentum, and energy \citep{Hawley84, Norman86}.  A variety of
Riemann solvers have been implemented.

@@

\paragraph{Godunov MHD with Constrained Transport (MHD)}
This MHD method is second-order in time and space, and preserves the
divergence constraint, $\div \vecB = 0$, to machine precision
through the Constrained Transport (CT) method \citep{Collins10}.  CT, originally
described by \citet{Evans88}, updates the magnetic field with the curl
of an electric field, suitably formulated to preserve $\div
\vecB$.  We employ the CT methods described by \citet{Balsara99} and
\citet{Gardiner05} with the second-order hyperbolic
solver of \citet{Li08a} and the divergence free AMR scheme of
\citet{Balsara01}.

\paragraph{Second-order finite difference method (HD only)}

Lastly, we briefly describe the \zeus\ method, a finite difference
algorithm originally used in the \zeus\ code \citep{Stone92a}. Note
that \enzo\ is entirely independent of the \zeus\ code, and only the
hydrodynamical algorithm of \zeus\ has been implemented in \enzo; the
MHD and radiation-hydrodynamics schemes have not. The \zeus\ method
uses a staggered mesh such that velocities are face-centered, while
density and internal energy are cell-centered.  It splits the solution
up into two steps. The first is the so-called source step, in which
the momentum and energy values are updated to reflect the pressure and
gravity forces, including an artificial viscosity required for
stability. The second step, known as the transport step, accounts for
the advection of conserved quantities (mass, momentum, and energy)
across the grid.
